---
title: "EDA v1"
author: "Marcus Needham"
date: "2024-09-18"
output: html_document
---

# Introduction
This report provides an exploratory data analysis (EDA) of Swire Coca-Cola's machine breakdown and downtime data, with a focus on identifying key variables for building a predictive maintenance model. The analysis covers the exploration of machine conditions, parts involved in breakdowns, downtime durations, and trends over time.

# Libraries and Data Loading

```{r setup, include=FALSE}
# Load required libraries
library(tidyverse)  # For data manipulation and visualization
library(lubridate)  # For working with dates and times
library(ggplot2)    # For data visualization
```

```{r}
# Load the dataset
iwc_data <- read.csv("IWC_Work_Orders_Extract.csv")

# View the structure of the dataset
str(iwc_data)
head(iwc_data)
summary(iwc_data$ACTUAL_WORK_IN_MINUTES)
```

```{r}
# Remove rows with missing values in ACTUAL_WORK_IN_MINUTES
iwc_data_clean <- iwc_data %>%
  filter(!is.na(ACTUAL_WORK_IN_MINUTES))

# Plot the distribution after cleaning data
ggplot(iwc_data_clean, aes(x = ACTUAL_WORK_IN_MINUTES)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Actual Work in Minutes", 
       x = "Actual Work (minutes)", 
       y = "Frequency") +
  theme_minimal()
```

This plot shows the overall distribution of actual work times. Most maintenance events fall within a lower range of time, with few outliers at higher durations.

```{r}
# Remove extreme outliers for better visualization (e.g., work longer than 1000 minutes)
iwc_data_filtered <- iwc_data_clean %>%
  filter(ACTUAL_WORK_IN_MINUTES <= 1000)  # Adjust threshold based on data distribution

# Plot after filtering outliers
ggplot(iwc_data_filtered, aes(x = ACTUAL_WORK_IN_MINUTES)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Actual Work in Minutes", 
       x = "Actual Work (minutes)", 
       y = "Frequency") +
  theme_minimal()
```

Here is the distribution of Actual Work time in minutes. You can see that most of the work on maintenance is below 200 minutes.

```{r}
# Group by PLANT_ID and count the number of downtimes per plant
plant_downtime_counts <- iwc_data %>%
  group_by(PLANT_ID) %>%
  summarise(downtime_count = n())

# Plot the downtime frequency per plant
ggplot(plant_downtime_counts, aes(x = reorder(PLANT_ID, -downtime_count), y = downtime_count)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Frequency of Downtimes by Plant", 
       x = "Plant ID", 
       y = "Number of Downtimes") +
  theme_minimal() +
  coord_flip()  # Flip the coordinates if the plant names are too long
```

This bar chart identifies which plants experience the highest frequency of downtime, highlighting potential areas for operational improvement.

### Looking at common causes of downtime

```{r}
# Group by maintenance description and count occurrences
common_causes <- iwc_data %>%
  group_by(ORDER_DESCRIPTION, MAINTENANCE_TYPE_DESCRIPTION) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# View the top 10 most common causes (all are over 1000)
head(common_causes, 12)
```

This the most common causes of maintenance. DTRINE is the highest with over 2,000 cases but the rest are all above 1,000 cases.

```{r}
common_maintenance_types <- iwc_data %>%
  group_by(MAINTENANCE_TYPE_DESCRIPTION) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# View the most common maintenance types
print(common_maintenance_types)
```

This shows whether the maintenance was preventative or corrective as most of them are either. 

```{r}
# Count unplanned vs. planned downtimes
planned_vs_unplanned <- iwc_data %>%
  group_by(MAINTENANCE_ACTIVITY_TYPE) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(count))

# View the breakdown of unplanned vs. planned
print(planned_vs_unplanned)
```

Unplanned maintenance happens almost 90% of the time over planned maintenance. 

### Time to resolve downtimes

```{r}
# Calculate mean and median of downtime per location
avg_downtime_by_location <- iwc_data %>%
  group_by(PLANT_ID) %>%
  summarize(
    mean_downtime = mean(ACTUAL_WORK_IN_MINUTES, na.rm = TRUE),
    median_downtime = median(ACTUAL_WORK_IN_MINUTES, na.rm = TRUE),
    count = n()
  ) %>%
  arrange(desc(mean_downtime))

# View the result
print(avg_downtime_by_location)

library(ggplot2)

# Plot average downtime by location
ggplot(avg_downtime_by_location, aes(x = reorder(PLANT_ID, -mean_downtime), y = mean_downtime)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(
    title = "Average Downtime Duration by Location",
    x = "Plant ID",
    y = "Mean Downtime (Minutes)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This visualization highlights which locations experience the highest average downtime, enabling targeted interventions to reduce delays. G221 with the most.

```{r}
# Calculate mean and median downtime by maintenance type
downtime_by_maintenance_type <- iwc_data %>%
  group_by(MAINTENANCE_TYPE_DESCRIPTION) %>%
  summarize(
    mean_downtime = mean(ACTUAL_WORK_IN_MINUTES, na.rm = TRUE),
    median_downtime = median(ACTUAL_WORK_IN_MINUTES, na.rm = TRUE),
    count = n()
  ) %>%
  arrange(desc(mean_downtime))

# View the result
print(downtime_by_maintenance_type)

# Plot downtime by maintenance type
ggplot(downtime_by_maintenance_type, aes(x = reorder(MAINTENANCE_TYPE_DESCRIPTION, -mean_downtime), y = mean_downtime)) +
  geom_bar(stat = "identity", fill = "lightcoral") +
  theme_minimal() +
  labs(
    title = "Average Downtime Duration by Maintenance Type",
    x = "Maintenance Type",
    y = "Mean Downtime (Minutes)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Maintenance types with higher average downtime durations represent key areas for process improvement to enhance operational efficiency. Breakdown Maintenance Order is the highest average downtime duration.

